---
title: "screenr Demonstration"
author:
- name: Steve Gutreuter
  email: sgutreuter@gmail.com
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{screenr Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
library(screenr)
```

# Introduction {-}

The screenr package enables easy development and validation of test
diagnostic screening tools of the sort reviewed by Clemens et al
[1].  This vignette demonstrates the development and
validation of a test-screening tool using lasso-like *L*1 regularization
[2] of logistic regression, as implemented in the glmpath
package [3].  The screenr package also supports
maximum-likelihood estimation of logistic regression parameters, but
the regularization approach is simpler to use, prevents overfitting
and handles separation in logistic regression.  Model performance is evaluating using the
receiver-operating characteristics [4,5].

## Training Data {-}

A sample of `r dim(unicorns)[1]` properly consented [unicorns](https://www.britannica.com/topic/unicorn) were enrolled in a study
aimed at improving testing for Unicorn
Immunodeficiency Virus (UIV).  There is currently no cure for UIV infection, but
highly effective antiretrovial therapy has the potential to both block
forward transmission and to avert death from opportunistic infections
associated with UIV infection.  Therefore is critically important that
UIV infection is detected and that infected individuals are enrolled
in treatment.  The subjects were administered a questionnaire
idenfying the presence or absence of putative risk factors for UIV
infection.  The prevalence of UIV is low, and therefore universal
testing was deemed impractical.  The challenge then is to identify
unicorns who should be prioritized for testing. Because UIV is
transmissible and fatal if untreated, it is important that the
screening tool have an acceptably high sensitivity.  The screenr
package enables development and validation of such screening tools.

The screening questionnaire included seven questions, which were
selected based on pubished information on infection risk.  The data
consist of the responses to the screening questions, coded as 1 for an
affirmative (yes) response and 0 for a negative response, and the test
result, again coded 0 and 1 for negative and positive results,
respectively.

```{r}
## The first six lines of the unicorns data:
head(unicorns)
```
It is critically important that all of the putative risk factors have
a positive association with the outcome (testresult in the
unicorns data).  That is, the questionnaire must be designed so that
affirmative (yes) reponses are hypthesized to indicate an *increased*
risk of UIV infection.

# Screening Tool Development {-}

Tool development consists of as many as four steps:
\describe{
\item Model fitting
\item Assessment of model performance based on the training data
\item Validation on new data
\item Evaluation of simplified implementation.
}

## Model fitting

The first step is estimation of the logistic-regression parameters
from the training data.  The function glmpathScreenr provides easy
integrated access to the glmpath [5] and pROC [6] packages using a convenient
formula interface.

```{r}
uniobj1 <- glmpathScreenr(testresult ~ Q1 + Q2 + Q3 + Q4 + Q5 + Q6,
                          data = unicorns, Nfolds = 10)
class(uniobj1)
```

The resulting glmpathScreer-class object includes the results from
regularization and receiver-operating characteristics (ROCs) from the
fits to all of the data, and from *k*-fold cross validation.  In the
above call, out-of-sample performance was estimated using 10 folds
(partitions of the data).  The resulting object, uniobj1, contains all
of the information need to develop and validate a screening tool for
diagnostic tests.

The GLM path-following algorithm [3] computes the
logit-scale coefficients along the entire regularization path, and has
the effect of selecting covariates based on penalization of *L*1-norm
of their coefficients, similar to the lasso [2].  Unlike
the lasso, the Park-Hastie algorithm also imposes a very small penalty
on the *L*2-norm, which eliminates adverse effects of any strong
correlations among the predictors and provides useful estimates even
if the response variable is separable by the predictors.  Two important
solutions along the path are given special status in glmpathScreenr
objects.  The solutions which yielded the smallest AIC and BIC are
denoted "minAIC" and "minBIC", respectively.  Those are likely the
most useful solutions in nearly all settings.  The logit-scale
coefficients for those solutions are obtained using the following
code:

```{r}
coef(uniobj1)
```
Note that the coefficient for Q3 has been shrunk to zero in both the
AIC- and BIC-best models which, in this case, happen to coincide.  In
effect, Q3 has been eliminated from both of those models. Q3 was not
an important predictor of the test result.

One can also obtain the odds ratios and/or omit the intercept.  For
example, the adjusted odds ratios for the predictors can be obtained
using:

<!-- ```{r} -->
<!-- coef(uniobj1, or = TRUE, intercept = FALSE) -->
<!-- ``` -->
One can also examine the coefficients everywhere the active set
changed along the regularization path:

<!-- ```{r} -->
<!-- pathobj <- getWhat(from = uniobj1, what = "glmpathObj", model = "minAIC") -->
<!-- plot(pathobj) -->
<!-- ``` -->

## Performance evaluation {-}

The next step is selection of a screening threshold based on the
ROC.  The ROC from the solution which minimized AIC is obtained using,

<!-- ```{r} -->
<!-- plot(uniobj1, print = TRUE) -->
<!-- ``` -->
which plots the overly-optimistic in-sample ROC and the
cross-validated estimate of out-of-sample performance, and optionally
returns a dataframe containing thresholds and sensitivities and
specificities along with their bootstrap confidence intervals.  The
cross-validated ROC provides options for selection of screening thresholds.

# References {-}

1. Clemens SL, Macneal KD, Alons CL, Cohn JE. Screening algorithms to
reduce burden of pediatric HIV testing: A systematic review and
meta-analysis. The Pediatric Infectious Disease Journal. 2020;
39(10):e303-e309. DOI: 10.1097/INF.0000000000002715.

2. Hastie T, Tibshirani R, Friedman R. The Elements of Statistical
   Learning. Springer, New York. 2009.  DOI: 10.1007/b94608.

3. Park MY, Hastie T. *L*1-regularization path algorithm for
   generalized linear models. Journal of the Royal Statistical Society
   Series B. 2007; 69(4):659-677.

4. Linden A. Measuring diagnostic and predictive accuracy in disease
   management: an introduction to receiver operating characteristic
   (ROC) analysis. Journal of Evaluation in Clinical Practice. 2006;
   12(2):132-139. DOI: 10.1111/j.1365-2753.2005.00598.x

5. Fawcett T. An introduction to ROC analysis.  Pattern Recognition
   Letters. 2006; 27(8):861-874. DOI: 10.1016/j.patrec.2005.10.010.

6. Robin X, Turck N, Hainard A, Tiberti N, Lisacek F, Sanchez J-C,
   Müller M. pROC: an open-source package for R and S+ to analyze and
   compare ROC curves. BMC Bioinformatics. 2011; 12(77):1-8. DOI:
   10.1186/1471-2105-12-77.
